---
title: "Homework 04: Data frames and data wrangling"
format: html
editor: visual
---

Please read the entire chapter on data transformation from\
[R for Data Science](https://r4ds.hadley.nz/data-transform) before starting this homework.

This homework relies on the `nycflights13` package, which contains several data frames, including:

-   `airlines`
-   `airports`
-   `flights`
-   `planes`
-   `weather`

Loading `nycflights13` puts these data frames on the search path.

------------------------------------------------------------------------

## Setup

### Load packages (do this every time)

```{r}
library(tidyverse)
library(nycflights13)
```

Note - if you don't have these installed you will need to first install.packages('nycflights13')

------------------------------------------------------------------------

# Question 1: Filtering

Make a plot of **air time vs distance** (air time on the y-axis, distance on the x-axis) for all flights that meet the following criteria:

-   originate from LaGuardia airport (`"LGA"`)
-   departed on the 16th of the month
-   have a flight distance of less than 2000 miles

## Your code

```{r}
flights |>
  filter(distance < 2000) |>
  filter(origin == "LGA") |>
  ggplot(
    mapping = aes(x=distance, y=air_time)
  ) +
  geom_point(na.rm = TRUE)
  


```

## Brief written response

In 2–4 sentences, describe what you observe in the plot. (For example: Is the relationship roughly linear? Are there any clear outliers?)

**Your answer here:**

the airtime is surprisingly linear, on average. however, it appears that the slope would not be one, if an analysis were to be done. Travel includes less air time per mile for every mile added to the distance of the flight. This makes sense considering the time needed for take off and descent. I don' tsee major outliers, however it is interesting that there is a pretty big gap in records between 1200-1400 mile flights.

------------------------------------------------------------------------

# Question 2: Dealing with NAs

Make a data frame of all rows of `flights` that have values for *both* `arr_time` and `dep_time` (i.e., neither value is `NA`).

## Your code

```{r}
# Create a new data frame that removes rows where arr_time or dep_time is NA.
Q2flights <- flights |>
  filter(!is.na(arr_time), !is.na(dep_time))
View(Q2flights)
```

## Filtering NAs (conceptual)

`ggplot()` will automatically remove `NA` values from a plot, but it emits a warning message about it. You *could* silence warnings using chunk options, but instead:

### Brief written response

Explain (in words) how you could prevent those `NA` values from appearing in the plot in the first place.

**Your answer here: using what I used in Q1, (na.rm = TRUE) removes these NA values before performing the task. I could also do what we just did in Q2 and filter out NAs myself**

------------------------------------------------------------------------

# Question 3: Adding columns

Create a data frame of **average flight speeds**, based on `air_time` and `distance`.

Then make either:

-   a histogram, **or**
-   a density plot

If you like, you may break the data out (e.g., by airline or another variable) in a way that you think makes sense.

## Your code

```{r}
# Create a new column for average speed.
# (Hint: think carefully about units — air_time is in minutes.)
# Then make a histogram or density plot.

Q3flights <- flights |>
  mutate(speed_mph = distance/air_time *60)

Q3flights |>
  ggplot(aes(x= speed_mph))+
  geom_histogram()

```

## Brief written response

Describe the main features of your speed distribution.

**Your answer here: flights tend to move at 400mph. and don't range too far away from 400mph. Fastest speed reached is just over 500, and oly a flight or two really drifts to 100mph**

------------------------------------------------------------------------

# Rendering and submission (GitHub)

Canvas contains the course GitHub link and instructions for **forking** the course repository.

In this file, your job is to:

1.  Complete the homework by adding code + written responses above.
2.  Render this Quarto file to **HTML or PDF**.
3.  Move the rendered file(s) to the /doc folder (note that, if you render to html you will also have to move your hmk_04_data_frames_files folder)
4.  Add/commit both:
    -   this edited `.qmd` file
    -   the rendered output file(s) (HTML or PDF)
5.  Push your changes to your fork on GitHub.
6.  Open a Pull Request back to the course repository.

## What to submit on Canvas

On Canvas, submit a link to your GitHub fork repository (not files).
